MSDS_6372_Project2

### 2022-07-29 MSDS_6372_Project2.Rmd
more progress

```{r, vif function}
# VIF score function
vif_score = function(model_name)
{
  vif_values = vif(model_name)[,3]
  barplot(vif_values, main = 'VIF Values', horiz = TRUE, col="blue", xlim = c(0,12))
  abline(v=10, col="red")
  return (vif_values)
}
```

@@

################ LASSO ROC Curve ######################################

lasso.roc.test = prediction(lasso.mod.final.pred.test, test$fracture,label.ordering=c("No","Yes"))
roc.lasso.test = performance(lasso.roc.test, measure = "tpr", x.measure = "fpr")
plot(roc.lasso.test,colorize = TRUE)
abline(a=0, b= 1)

AUC = auc(test$fracture, lasso.mod.final.pred.test)
print(paste("Area Under the Cuve: ", AUC))
print(paste("LASSO AIC Score: ", AIC(lasso.mod.final)))

#alternative approach
lasso.roc<-roc(response=test$fracture,predictor=lasso.mod.final.pred.test,levels=c("No","Yes"))
plot(lasso.roc,print.thres="best") #This graph is nice because the x axis is plotted in terms of specificity rather than FPR
auc(lasso.roc)

cutoff.lasso = 0.212

@@

#Confusion Matrix for LASSO
print("Confusion matrix for LASSO TRAINING")
confusionMatrix(table(class.lasso.final.train,train$fracture), positive = "Yes")

################ Misclassification rate train ######################################
#cross.table.lasso.train = table(table(class.lasso.final.train,train$fracture))
#MCR_lasso.train = (cross.table.lasso.train[2]+cross.table.lasso.train[3])/dim(train)[1]
misClasificError.lasso.train = mean(class.lasso.final.train != train$fracture)
print(paste('Misclassification Rate for lasso on training set: ', misClasificError.lasso.train))
#####################################################################################################

# Confusion matrix test
class.lasso.final.test<-factor(ifelse(lasso.mod.final.pred.test>cutoff.lasso,"Yes","No"),levels=c("No","Yes"))

#Confusion Matrix for LASSO
print("Confusion matrix for LASSO TEST with 0.5 cutoff")
confusionMatrix(table(class.lasso.final.test,test$fracture), positive = "Yes")

################ Misclassification rate test ######################################
#cross.table.lasso.test = table(class.lasso.final.test,test$fracture)
#MCR_lasso.test = (cross.table.lasso.test[2]+cross.table.lasso.test[3])/dim(test)[1]
#print(paste('Misclassification Rate for LASSO selection on test set: ', MCR_lasso.test))
misClasificError.lasso.test = mean(class.lasso.final.test != test$fracture)
print(paste('Misclassification Rate for lasso on test set: ', misClasificError.lasso.test))
```

## Assumprions
```{r, fit after feature selection2}
# Checking logistic regression model assumptions

# The Observations are Independent
# We can assume that the observations are not coming from repeated measures

#There is No Multicollinearity Among Explanatory Variables
### Visualize VIF
#vif_score(lasso.mod)

# There are No Extreme Outliers (cooks D)
#influenceIndexPlot(lasso.mod)

# There is a Linear Relationship Between Explanatory Variables and the Logit of the Response Variable

# Influential point analysis and residual plots
#residualPlots(lasso.mod)
#influenceIndexPlot(lasso.mod)
#influencePlot(lasso.mod)
```

## Observations:
## All assumptions are met:
* __The Response Variable is Binary__: 'Fracture' as a response variable is a factor with binary levels (Yes/No)
* __Independence__: We can assume that the observations are independent
* __Multicolliearity__: The is no multicollinearity among the explanatory variables (VIF values show no multicollinearity)
* __Outliers__: The largest Cooks D value is 0.015 which indicates that there is no extreme outlier
* __The Sample Size is Sufficiently Large__: ?

* Sensitivity: The probability that the model predicts a positive outcome for an observation when indeed the outcome is positive. This is also called the “true positive rate.”
* Specificity: The probability that the model predicts a negative outcome for an observation when indeed the outcome is negative. This is also called the “true negative rate.”

## Model scoring
```{r model scoring lasso}
## Add misclassification rate
```

@@

@@

@@

### Visualize VIF
MLogReg_VIF = vif(MLogReg$finalModel)
barplot(MLogReg_VIF, main = 'VIF Values (Custom Logistic Regression', horiz = TRUE, col="blue", xlim = c(0,12))
abline(v=10, col="red")

@@

@@

@@

@@

lda.model = lda(fracture ~age+weight+height+bmi+fracscore, prior = c(.75, .25), data = lda.train)
plot(lda.model)

@@

@@

qda.model = qda(fracture ~age+weight+height+bmi+fracscore, prior = c(.75, .25), data = qda.train)
plot(lda.model)

@@

@@

@@

#Scoring
dt.prob.test = predict(dtree.model, dt.test, type = "prob")
AUC = auc(dt.test$fracture, dt.prob.test$Yes)
print(paste("Area Under the Cuve: ", AUC))
```

>>>>

```{r, vif function}
# VIF score function
vif_score = function(model_name)
{
  print(model_name)
  vif_values = vif(model_name)[,3]
  barplot(vif_values, main = 'VIF Values', horiz = TRUE, col="blue", xlim = c(0,12))
  abline(v=10, col="red")
  return (vif_values)
}
```

```{r, assumptions function}
# Assumptions function
assumptions = function(model_name)
{
  outliers = influenceIndexPlot(model_name)

# Influential point analysis and residual plots
  residual_plot = residualPlots(model_name)

  studentized = influencePlot(model_name)
  return ()
}
```

@@

################ LASSO ROC Curve ######################################
lasso.roc<-roc(response=test$fracture,predictor=lasso.mod.final.pred.test,levels=c("No","Yes"))
plot(lasso.roc,print.thres="best", main='LASSO ROC curve')

# Calculate Area Under the Curve and AIC score
AUC = auc(test$fracture, lasso.mod.final.pred.test)
print(paste("Area Under the Cuve: ", AUC))
print(paste("LASSO AIC Score: ", AIC(lasso.mod.final)))

#Set cut off value for labels
cutoff.lasso = 0.212

@@

#Confusion Matrix for LASSO - Train
print("Confusion matrix for LASSO TRAINING")
confusionMatrix(table(class.lasso.final.train,train$fracture), positive = "Yes")

#Confusion Matrix for LASSO - Test
print("Confusion matrix for LASSO TEST")
class.lasso.final.test<-factor(ifelse(lasso.mod.final.pred.test>cutoff.lasso,"Yes","No"),levels=c("No","Yes"))
confusionMatrix(table(class.lasso.final.test,test$fracture), positive = "Yes")

# Create variables to store metrics
lasso.final.test.accuracy = confusionMatrix(table(class.lasso.final.test,test$fracture), positive = "Yes")$overall[1]
lasso.final.test.sensitivity = confusionMatrix(table(class.lasso.final.test,test$fracture), positive = "Yes")$byClass[1]
lasso.final.test.specificity = confusionMatrix(table(class.lasso.final.test,test$fracture), positive = "Yes")$byClass[2]

##### Store model results for comparison ####
  Model_name = c('LASSO')
  AUC_score = c(AUC)
  Test_Accuracy  = c(lasso.final.test.accuracy)
  Test_Sensitivity  = c(lasso.final.test.sensitivity)
  Test_Specificity  = c(lasso.final.test.specificity)
  AIC  = c(AIC(lasso.mod.final))
  Threshold = c(cutoff.lasso)
```

## Logistic Regression Assumptions
```{r, assumtion logreg}
# Checking logistic regression model assumptions
vif_score(lasso.mod.final)
assumptions(lasso.mod.final)
```

## Observations:
## All assumptions are met:
* __The Response Variable is Binary__: 'Fracture' as a response variable is a factor with binary levels (Yes/No)
* __Independence__: We can assume that the observations are independent
* __Multicolliearity__: The is no multicollinearity among the explanatory variables (VIF values show no multicollinearity)
* __Outliers__: The largest Cooks D value is 0.02 which indicates that there is no extreme outlier
* __The Sample Size is Sufficiently Large__: ?

@@

# Create variables to store metrics
step.final.test.accuracy = confusionMatrix(table(class.stepwise.final.test,test$fracture), positive = "Yes")$overall[1]
step.final.test.sensitivity = confusionMatrix(table(class.stepwise.final.test,test$fracture), positive = "Yes")$byClass[1]
step.final.test.specificity = confusionMatrix(table(class.stepwise.final.test,test$fracture), positive = "Yes")$byClass[2]

@@

Model_name = c(Model_name, 'Stepwise')
AUC_score = c(AUC_score, AUC)
Test_Accuracy  = c(Test_Accuracy, step.final.test.accuracy)
Test_Sensitivity  = c(Test_Sensitivity, step.final.test.sensitivity)
Test_Specificity  = c(Test_Specificity, step.final.test.specificity)
AIC  = c(AIC, AIC(step.log))
Threshold = c(Threshold, cutoff.step)
```

## Logistic Regression Assumptions
```{r, fit after feature selection2}
# Checking logistic regression model assumptions
vif_score(step.log)
assumptions(step.log)
```

## Observations:
## All assumptions are met:
* __The Response Variable is Binary__: 'Fracture' as a response variable is a factor with binary levels (Yes/No)
* __Independence__: We can assume that the observations are independent
* __Multicolliearity__: The is no multicollinearity among the explanatory variables (VIF values show no multicollinearity)
* __Outliers__: The largest Cooks D value is 0.025 which indicates that there is no extreme outlier
* __The Sample Size is Sufficiently Large__: ?

@@

@@

# Create variables to store metrics
manual.final.test.accuracy = confusionMatrix(table(class.manual.test,test$fracture), positive = "Yes")$overall[1]
manual.final.test.sensitivity = confusionMatrix(table(class.manual.test,test$fracture), positive = "Yes")$byClass[1]
manual.final.test.specificity = confusionMatrix(table(class.manual.test,test$fracture), positive = "Yes")$byClass[2]

Model_name = c(Model_name, 'Manual/Intuition model')
AUC_score = c(AUC_score, AUC)
Test_Accuracy  = c(Test_Accuracy, manual.final.test.accuracy)
Test_Sensitivity  = c(Test_Sensitivity, manual.final.test.sensitivity)
Test_Specificity  = c(Test_Specificity, manual.final.test.specificity)
AIC  = c(AIC, MLogReg$finalModel$aic)
Threshold = c(Threshold, cutoff.manual)
```

## Logistic Regression Assumptions
```{r, assumption logreg}
### Visualize VIF
MLogReg_VIF = vif(MLogReg$finalModel)
barplot(MLogReg_VIF, main = 'VIF Values (Custom Logistic Regression', horiz = TRUE, col="blue", xlim = c(0,12))
abline(v=10, col="red")
```

## Observations:
## All assumptions are met:
* __The Response Variable is Binary__: 'Fracture' as a response variable is a factor with binary levels (Yes/No)
* __Independence__: We can assume that the observations are independent
* __Multicolliearity__: The is no multicollinearity among the explanatory variables (VIF values show no multicollinearity)
* __Outliers__: The largest Cooks D value is 0.02 which indicates that there is no extreme outlier
* __The Sample Size is Sufficiently Large__: ?

@@

# Create variables to store metrics
MLogReg.ii.final.test.accuracy = confusionMatrix(table(class.manual.test.ii,test$fracture), positive = "Yes")$overall[1]
MLogReg.ii.final.test.sensitivity = confusionMatrix(table(class.manual.test.ii,test$fracture), positive = "Yes")$byClass[1]
MLogReg.ii.final.test.specificity = confusionMatrix(table(class.manual.test.ii,test$fracture), positive = "Yes")$byClass[2]

Model_name = c(Model_name, 'Complex model')
AUC_score = c(AUC_score, AUC)
Test_Accuracy  = c(Test_Accuracy, MLogReg.ii.final.test.accuracy)
Test_Sensitivity  = c(Test_Sensitivity, MLogReg.ii.final.test.sensitivity)
Test_Specificity  = c(Test_Specificity, MLogReg.ii.final.test.specificity)
AIC  = c(AIC, MLogReg.ii$finalModel$aic)
Threshold = c(Threshold, cutoff.manual)

@@

```
## Logistic Regression Assumptions
```{r, assumptions}
### Visualize VIF
MLogReg.ii_VIF = vif(MLogReg.ii$finalModel)
barplot(MLogReg_VIF, main = 'VIF Values (Custom Logistic Regression', horiz = TRUE, col="blue", xlim = c(0,12))
abline(v=10, col="red")
```

## Observations:
## All assumptions are met:
* __The Response Variable is Binary__: 'Fracture' as a response variable is a factor with binary levels (Yes/No)
* __Independence__: We can assume that the observations are independent
* __Multicolliearity__: The is no multicollinearity among the explanatory variables (VIF values show no multicollinearity)
* __Outliers__: The largest Cooks D value is 0.02 which indicates that there is no extreme outlier
* __The Sample Size is Sufficiently Large__: ?

@@

lda.model = lda(fracture ~age+weight+height, prior = c(.75, .25), data = lda.train)
plot(lda.model)

@@

# Create variables to store metrics
lda.final.test.accuracy = confusionMatrix(table(lda.pred.adj.test,test$fracture), positive = "Yes")$overall[1]
lda.final.test.sensitivity = confusionMatrix(table(lda.pred.adj.test,test$fracture), positive = "Yes")$byClass[1]
lda.final.test.specificity = confusionMatrix(table(lda.pred.adj.test,test$fracture), positive = "Yes")$byClass[2]

Model_name = c(Model_name, 'LDA')
AUC_score = c(AUC_score, AUC)
Test_Accuracy  = c(Test_Accuracy, lda.final.test.accuracy)
Test_Sensitivity  = c(Test_Sensitivity, lda.final.test.sensitivity)
Test_Specificity  = c(Test_Specificity, lda.final.test.specificity)
AIC  = c(AIC, 'N/A')
Threshold = c(Threshold, cutoff.lda)

@@

qda.model = qda(fracture ~age+weight+height, prior = c(.75, .25), data = qda.train)
plot(lda.model)

@@

# Create variables to store metrics
qda.final.test.accuracy = confusionMatrix(table(qda.pred.adj.test,test$fracture), positive = "Yes")$overall[1]
qda.final.test.sensitivity = confusionMatrix(table(qda.pred.adj.test,test$fracture), positive = "Yes")$byClass[1]
qda.final.test.specificity = confusionMatrix(table(qda.pred.adj.test,test$fracture), positive = "Yes")$byClass[2]

Model_name = c(Model_name, 'QDA')
AUC_score = c(AUC_score, AUC)
Test_Accuracy  = c(Test_Accuracy, qda.final.test.accuracy)
Test_Sensitivity  = c(Test_Sensitivity, qda.final.test.sensitivity)
Test_Specificity  = c(Test_Specificity, qda.final.test.specificity)
AIC  = c(AIC, 'N/A')
Threshold = c(Threshold, cutoff.qda)

@@

## LDA/QDA Assumptions
```{r, lda qda assumptions}
### Visualize VIF
MLogReg.ii_VIF = vif(MLogReg.ii$finalModel)
barplot(MLogReg_VIF, main = 'VIF Values (Custom Logistic Regression', horiz = TRUE, col="blue", xlim = c(0,12))
abline(v=10, col="red")
```

## Observations:
* __Multivariate normality__: 
* __Equal covariance matrices__:

@@

#Scoring
dt.prob.test = predict(dtree.model, dt.test, type = "prob")
AUC = auc(dt.test$fracture, dt.prob.test$Yes)
print(paste("Area Under the Cuve: ", AUC))

# Create variables to store metrics
dtree.final.test.accuracy = confusionMatrix(table(dtree.pred.adj.test,test$fracture), positive = "Yes")$overall[1]
dtree.final.test.sensitivity = confusionMatrix(table(dtree.pred.adj.test,test$fracture), positive = "Yes")$byClass[1]
dtree.final.test.specificity = confusionMatrix(table(dtree.pred.adj.test,test$fracture), positive = "Yes")$byClass[2]

Model_name = c(Model_name, 'Decision Tree')
AUC_score = c(AUC_score, AUC)
Test_Accuracy  = c(Test_Accuracy, dtree.final.test.accuracy)
Test_Sensitivity  = c(Test_Sensitivity, dtree.final.test.sensitivity)
Test_Specificity  = c(Test_Specificity, dtree.final.test.specificity)
AIC  = c(AIC, 'N/A')
Threshold = c(Threshold, cutoff.dtree)

```{r, model summary}
# model summary
model.sum.df = data.frame()
model.sum.df = data.frame(Model_name, AUC_score, Test_Accuracy, Test_Sensitivity, Test_Specificity, AIC, Threshold)
model.sum.df
